%\documentclass{acm_proc_article-sp}
%\documentclass{sig-alternate}
\documentclass[preprint, 10pt, numbers]{sigplanconf}

\usepackage{tikz}         % for graphics
\usepackage{caption}
\usepackage{pgfplots}
\usepackage{amssymb}      % some useful maths symbols
\usepackage{pifont}       % to get dingbats
\usepackage{url}          % for typesetting URLs
\usepackage{fancyvrb}     % for code snippets
\usepackage{listings}
\usepackage{paralist}
\usepackage{color}                % only for autor notes to each other
%\def\comment#1{{\color{red}#1}}  % display comments/notes.
%\def\comment#1{{\color{red}}}    % hide comments/notes.

\lstset{language=C, basicstyle=\ttfamily, xleftmargin=2pc}
\lstset{
  literate={\_}{}{0\discretionary{\_}{}{\_}}%
}
\newcommand{\co}[1]{\lstinline[breaklines=true,breakatwhitespace=true]{#1}}
\newcommand{\mkcol}[1]{\textcolor{red}{\textbf{#1}}}

\begin{document}

% ---------------------------------------------------------------------
% Title and authors,  adherence to SIGS style
% ---------------------------------------------------------------------
\title{Verification of the Tree-Based Hierarchical Read-Copy~Update in the Linux Kernel}
\date{}

% sigplan template
%\authorinfo{\textit{Authors and affiliations omitted, for double-blind review.}}

\authorinfo{Lihao Liang}
           {University of Oxford}
           {lihao.liang@cs.ox.ac.uk}

\authorinfo{Paul E. McKenney}
           {Linux Technology Center, IBM}
           {paulmck@linux.vnet.ibm.com}

\authorinfo{Daniel Kroening}
           {University of Oxford}
           {daniel.kroening@cs.ox.ac.uk}

\authorinfo{Tom Melham}
           {University of Oxford}
           {tom.melham@cs.ox.ac.uk}

% sig-alternate template
%\numberofauthors{4}
%\author{
%%\textit{Authors and affiliations omitted, for double-blind review.}
%%
%  Lihao Liang \\
%  \affaddr{University of Oxford} \\
%  %\email{lihao.liang@cs.ox.ac.uk}
%  \alignauthor
%%
%  Paul E. McKenney \\
%  \affaddr{Linux Technology Center, IBM} \\
%  %\email{paulmck@linux.vnet.ibm.com}
%  \alignauthor
%%
%  Daniel Kroening \\
%  \affaddr{University of Oxford} \\
%  %\email{daniel.kroening@cs.ox.ac.uk}
%  \alignauthor
%%
%  Tom Melham \\
%  \affaddr{University of Oxford} \\
%  %\email{tom.melham@cs.ox.ac.uk}
%}

% ---------------------------------------------------------------------
% Start of text.
% ---------------------------------------------------------------------


\maketitle

\begin{abstract}
Read-Copy Update (RCU) is a scalable, high-performance Linux-kernel
synchronization mechanism that runs low-over\-head readers concurrently
with updaters.
Pro\-duc\-tion-quality RCU implementations for multi-core systems are
decidedly non-trivial. Giving the ubiquity of Linux, a rare
``million-year'' bug can occur several times per day across the installed
base.
Stringent validation of RCU's complex behaviors is thus critically important.
Exhaustive testing is infeasible due to the exponential
number of possible executions, which suggests use of formal verification.

Previous verification efforts on RCU either focus on simple implementations
or use modeling languages, the latter requiring
error-prone manual translation that must be repeated frequently
due to regular changes in the Linux kernel's RCU implementation.
In this paper, we first describe the implementation of Tree RCU in the
Linux kernel. We then discuss how to construct a model directly from Tree RCU's
source code in C, and use the CBMC model checker to verify its safety and liveness properties.
To our best knowledge, this is the first verification of a significant part
of RCU's source code, and is an important step towards integration of formal verification into
the Linux kernel's regression test suite.
\end{abstract}

\category{}{D.2.4}{Software/Program Verification}[Model checking]
\category{}{D.1.3}{Concurrent Programming}[Parallel programming]

\keywords
Software Verification, Parallel Computing, Read-Copy Update, Linux Kernel

% Include the sections of the paper.

\input{intro.tex}                 % Introduction
\input{background.tex}            % Background
\input{tree_rcu.tex}              % Tree RCU Implementation
\input{modeling.tex}              % Tree RCU Model
\input{experiments.tex}           % Experimental Studies
\input{related_work.tex}          % Related Work
\input{concl.tex}                 % Summary and Furture Work
%\input{ack.tex}                   % Acknowledgements

%\newpage
%\bibliographystyle{abbrv}    % sig-alternate
% \bibliographystyle{abbrvnat} % sigplanconf
% \raggedright
% \bibliography{paper}

%\clearpage
%\appendix
%\vskip5mm
%\input{appendix.tex}

\end{document}
